Distributional Trees and Forests

### 2017-11-23 pkg/disttree/inst/draft/draft.Rnw
small corrections
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1754 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 937236edecd689c8513921b0194a3773b8f4c0cc

%% packages
\usepackage{amstext}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{thumbpdf}
\usepackage{rotating}
\usepackage{caption}
\captionsetup{format=hang}
\usepackage{subcaption}
\usepackage{Sweave}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{array, makecell}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{positioning,shapes,arrows,decorations.pathreplacing,calc,automata,mindmap}
%\usepackage[pagebackref=true]{hyperref}
%\usepackage{doi}

@@

\Address{
  Lisa Schlosser \\
  Department of Statistics \\
  Faculty of Economics and Statistics \\
  Universit\"at Innsbruck \\
  Universit\"atsstr.~15 \\
  6020 Innsbruck, Austria \\
  E-mail: \email{Lisa.Schlosser@uibk.ac.at} \\
  URL: \url{https://www.uibk.ac.at/statistics/personal/schlosser-lisa/lisa-schlosser.html}\\

@@

  Reto Stauffer \\
  Department of Statistics \\
  Faculty of Economics and Statistics \\
  Universit\"at Innsbruck \\
  Universit\"atsstr.~15 \\
  6020 Innsbruck, Austria \\
  E-mail: \email{Reto.Stauffer@uibk.ac.at} \\
  URL: \url{https://www.uibk.ac.at/statistics/personal/stauffer/reto-stauffer.html}\\

  Achim Zeileis \\
  Department of Statistics \\
  Faculty of Economics and Statistics \\
  Universit\"at Innsbruck \\
  Universit\"atsstr.~15 \\
  6020 Innsbruck, Austria \\
  E-mail: \email{Achim.Zeileis@R-project.org} \\
  URL: \url{http://eeecon.uibk.ac.at/~zeileis/}

@@

After explaining the novel methodology in Section~\ref{Methodology} of this paper the focus will then be put on an application in the field of probabilistic forecasting in Section~\ref{ProbabilisticForecasting}. The new method is applied on precipitation data in a mountainous region with a large number of numerical weather prediction quantities and the results are compared to those of other methods commonly used in this research area. In a first setting only one observation station is taken into consideration. The models are trained on data from 24 years and probabilistic predictions of the consecutive 4 years are validated. A brief outlook on the results of this setting is given by Figure~\ref{axams_onestation}. For the 24th of July of 2009, 2010, 2011 and 2012 the density functions of the distributions estimated by a forest model are plotted together with the corresponding observations. This illustrates that well fitting predicted expected values are provided by this model as they are close to the observed values.
Beside the expected values the predictions provided by distributional models offer a wide range of information as they specify the whole distribution and therefore enable further methods of inference and analysis.

@@

## get predicted parameter
# for learndata and testdata
{
  ## disttree
  if(FALSE){
    # in sample
    pdt <- predict(res$dt, type = "parameter")
    dt_mu_l <- pdt$mu 
    dt_sigma_l <- pdt$sigma 
    pit_dt_l <- cbind(0, pnorm(res$learndata[,"robs"], mean = dt_mu_l, sd = dt_sigma_l))
    pit_dt_l[res$learndata[,"robs"]>0, 1] <- pit_dt_l[res$learndata[,"robs"]>0, 2]
  }  
  # out of sample
  pdt <- predict(res$dt, type = "parameter", newdata = res$testdata)
  dt_mu_t <- pdt$mu 
  dt_sigma_t <- pdt$sigma 
  pit_dt_t <- cbind(0, pnorm(res$testdata[,"robs"], mean = dt_mu_t, sd = dt_sigma_t))
  pit_dt_t[res$testdata[,"robs"]>0, 1] <- pit_dt_t[res$testdata[,"robs"]>0, 2]

@@

\begin{figure}[t!]
\begin{center}
<<rain_Axams_pred, fig=TRUE, echo=FALSE, width=6.5, height=4.5>>=
plot(x = x, y = y1, type = "l", col = "red", 
     main = paste0("July ", pday, dayending), ylab = "Density", 
     xlab = expression(Total~precipitation~"["~mm^(1/1.6)~"/"~"24h"~"]"),
     #xlab = expression(Total~precipitation~group("[", mm^(1/1.6)~per~day,"]")),
     #{\frac{1}{1.6}}$,
     ylim = c(0,max(y1, y2, y3, y4, pm1, pm2, pm3, pm4) + 0.01),
     xlim = c(-1.5,8))

@@

# add labels
text(x = -0.8, y = lh1, labels = "2009", col = "red", cex = 0.8)
text(x = -0.8, y = lh2, labels = "2010", col = "blue", cex = 0.8)
text(x = -0.8, y = lh3, labels = "2011", col = "darkgreen", cex = 0.8)
text(x = -0.8, y = lh4, labels = "2012", col = "purple", cex = 0.8)
@
\caption{\label{axams_onestation}Predictions at station Axams for one day in July of 2009, 2010, 2011 and 2012. Observations and predictions are left censored at 0. The corresponding point mass is shown at the censoring point (0).}
\end{center}
\end{figure}

\newpage
\section{Methodology}
\label{Methodology}

@@

In our application of precipitation forecasting the response variable, which is the power transformed total precipitation amount within 24 hours, is expected to be normally distributed with mean/location parameter $\mu$ and standard deviation/scale parameter $\sigma$ and left censored with censoring point 0. (The power transformation is used to remove large parts of the skewness of the data as further explained in \ref{setting}.) Therefore, the corresponding log-likelihood function with parameter vector $\theta = (\mu, \sigma)$ is

@@

The exact way of determining how and where to split varies across different tree algorithms. (For distributional trees, one of two algorithms can be chosen as a framework to build the tree: either the MOB algorithm or the ctree algorithm.) Summed up, the steps of building a distributional tree using one of these algorithms are
\begin{enumerate}
\item Specify a distribution with log-likelihood function $\ell(\theta; y)$.
\item Estimate $\bold{\hat{\theta}}$ via maximum likelihood.
\item Test for associations or instabilities of the scores $s(\bold{\hat{\theta}}, y_i)$ and each partitioning variable $Z_l$.
%\item Assess whether the \emph{model scores} are associated with
%    (or change along) any of the available covariates -- e.g.,
%    using parameter instability tests (\emph{strucchange}) or
%    conditional inference (\emph{coin}).
\item Split the sample along the partitioning variable with the strongest association or instability.
    Choose the breakpoint with the highest improvement in log-likelihood.
\item Repeat steps 2--4 recursively in the subgroups
    until some stopping criterion is reached.
    %-- e.g., for significance or sample size.
%\item Choose the variable with the strongest association
%\item Choose the split point which leads to the highest improvement of the model
%\item Split and repeat 2-6 in each node until a stopping criterion is met
\end{enumerate}
The stopping criterion in step 5 can for example be a minimum number of observations in a node or a p-value for the statistical tests applied in step 3.\\
Executing these steps results in a stepwise parametric model where a complete distribution is fit in each node of the tree.\\
\\
For distributional trees the strategy of how to make a prediction for a new set of covariates $\bold{z}^{\ast} = (z_1^{\ast}, \ldots, z_m^{\ast})$ is very simple and straight forward. All observations of the learning data set $\{(y_i,\bold{z}_i)\}_{i=1\ldots,n}$ which end up in the same terminal node of the tree as the new set of covariates are selected and a distributional model is fit on the subset. To select this subset a weight vector $\bold{\omega}^{tree} = (\omega^{tree}_1, \ldots, \omega^{tree}_n)$ is included in the fitting process.
$$
\omega^{tree}_i(\bold{z}^{\ast}) = \sum_{b=1}^B I((\bold{z}_i \in \mathcal{B}_b) \land (\bold{z}^{\ast} \in \mathcal{B}_b))
$$
The predicted distribution of the new set of covariates is then fully specified by the estimated parameter $\hat{\theta}(\bold{z}^{\ast})$ where
$$
\hat{\theta}(\bold{z}^{\ast}) = \max_{\theta \in \Theta} \sum_{i=1}^n \omega^{tree}_i(\bold{z}^{\ast}) \cdot \ell(\theta; y_i).
$$

@@

\SweaveOpts{eval = TRUE}
\newpage
\newpage
\section{Probabilistic Forecasting of Precipitation}
\label{ProbabilisticForecasting}

@@

\subsection{Setting}
\label{setting}
The data set used for training and validation consists of observed daily precipitation sums provided by the National Hydrographical Service (\citealt{ehyd}) and numerical weather forecasts from the {U.S.} National Oceanic and Atmospheric Administration (NOAA).
Both, observations and weather forecasts are available from 1985 trough the year 2012, focusing on July only in this article.\\ 
Observations are available on a daily basis, each observation representing the total precipitation sum fallen within 24~hours observed at 6~UTC at 95 stations all over Tyrol and it's close surrounding. The observations are undergoing a rigorous quality check before made available to the public.  As covariates the second generation reforecast data set of the global ensemble forecast system (GEFS) is used (\citealt{hamill2013}).  This data set consists of an 11-member ensemble based on a fixed version of the numerical model and a horizontal grid-spacing of about $50~\times~50 \text{km}^2$ initialized at 0~UTC from December 1984 to present. Each of the 11 ensemble members uses slightly different perturbed initial conditions trying to predict the situation specific uncertainty of the atmospheric state.

@@

In particular, 14 basic forecast variables are considered with up to 12 variations of each of these variables being included as covariates. Since each forecast variable consists of an ensemble of individual predictions, the possible variations include the mean, the standard deviation, the minimum and the maximum over all ensemble members, noted as ensemble mean, ensemble std dev, ensemble minimum and ensemble maximum. All of these values can be calculated on the sums over 24 hours (6--30 UTC) or a 6 hour time window.\\
Another possible option is to consider the mean, minimum or maximum over 24 hours for each individual prediction and then take the ensemble mean or ensemble standard deviation, noted as ensemble mean of mean/minimal/maximal values or ensemble std dev of mean/minimal/maximal values respectively.\\
Table~\ref{covariates} shows all variables together with the number of variations and the provided variations listed in the last column. Altogether 82 covariates are available in this setting.\\
To remove large parts of the skewness of precipitation data, a power transformation has been used frequently (Box and Cox 1964). In literature, cubic (\citealt{stidd1973}) or square root (\citealt{hutchinson1998b}) transformations have often been suggested but may vary for different climatic zones or temporal aggregation periods. In \citealt{stauffer2017a} optimization methods have been applied to find the best transformation factor which resulted in a value of $\frac{1}{1.6}$. Since the data used for this optimization covers almost the same geographical region as considered in this article the same factor is chosen here for the power transformation of the response variable \texttt{tp} (total precipitation). The resulting transformed variable is expected to be normally distributed and left censored at censoring point 0. For simplicity the same transformation has been chosen for the variable \texttt{cape} (convective available potential energy).

%\newpage
\begin{table}[t!]
%\resizebox{1.4\textwidth}{!}{
\begin{minipage}{\textwidth}
\begin{tabular}{ l  c  l }
\hline
Basic covariates & {\#} & Variations\\
\hline
\textbf{tp}: total precipitation            & 12 variations:  & ensemble mean of sums over 24h, \\
\hspace*{\fill} power transformed (by $\frac{1}{1.6}$) &    & ensemble std dev of sums over 24h, \\ 
\textbf{cape}: convective                  &    & ensemble minimum of sums over 24h, \\
\hspace*{0.4cm} available potential energy &    & ensemble maximum of sums over 24h\\
\hspace*{\fill}  power transformed (by $\frac{1}{1.6}$)&    & \qquad all for 6--30 UTC\\
\hspace*{\fill}                            &    & ensemble mean of sums over 6h\\
\hspace*{\fill}                            &    & \qquad for 6--12, 12--18, 18--24, 24--30 UTC\\
\hspace*{\fill}                            &    & ensemble std dev of sums over 6h\\
\hspace*{\fill}                            &    & \qquad for 6--12, 12--18, 18--24, 24--30 UTC\\
\hline
\textbf{dswrf}: downwards short ware      & 6 variations: & ensemble mean of mean values, \\
\hspace*{\fill} radiation flux (``sunshine'') &   & ensemble mean of minimal values,\\
\textbf{msl}: mean sea level pressure     &   &  ensemble mean of maximal values,\\
\textbf{pwat}: preciptable water          &   & ensemble std dev of mean values,\\
\textbf{tmax}: 2m maximum                 &   & ensemble std dev of minimal values,\\
\hspace*{1.15cm} temperature              &   & ensemble std dev of maximal values,\\
\textbf{tcolc}: total column-integrated   &   & \qquad all over 6--30 UTC \\
\hspace*{1.05cm} condensate  &   & \\
\textbf{t500}: temperature   &   & \\
\hspace*{1cm} on 500 hPa     &   & \\
\textbf{t700}: temperature   &   & \\
\hspace*{1cm} on 700 hPa     &   & \\
\textbf{t850}: temperature   &   & \\
\hspace{1cm} on 850 hPa      &   & \\
\hline
\textbf{tdiff500850}: temperature         & 3 variations: & ensemble mean of difference in mean,\\
\hspace*{\fill} difference 500 to 850 hPa &   & ensemble minimum of difference in mean,\\
\textbf{tdiff500700}: temperature         &   & ensemble maximum of difference in mean\\
\hspace*{\fill} difference 500 to 700 hPa &   & \qquad all over 6--30 UTC\\
\textbf{tdiff700850}: temperature         &   & \\
\hspace*{\fill} difference 700 to 850 hPa &   & \\
\hline
\textbf{msl{\_}diff}: mean sea level  & 1 variation: & msl{\_}mean{\_}max - msl{\_}mean{\_}min\\
\hspace{1.65cm} difference            &   & \qquad over 6--30 UTC\\
\hline
\end{tabular}
\caption[Table caption text]{Basic covariates together with the number and the type of variations.}
\label{covariates}
\end{minipage} 
%}
\end{table}

\subsection{Models and Evaluation}
Next to distributional forests three other methods are applied to this data set and their results are compared based on CRPS values. One of the reference models is an EMOS models as this is a common choice in this type of setting / research field. The other two models are both GAMs, once in a prespecified form in terms of variable selection for each distribution parameters and once applying a boosting algorithm). 
The chosen methods and their specifications are listed in Table~\ref{model_specification}.
\begin{table}[t!]
%\resizebox{1.4\textwidth}{!}{
%\begin{minipage}{\textwidth}
\begin{center}
\begin{tabular}{ l  l   l   l }
\hline
 & \textbf{Type} & \textbf{Location ($\boldsymbol{\mu}$)} & \textbf{Scale ($\boldsymbol{\log(\sigma)}$)} \\
\hline
\textbf{Forest} & recursive    & all & all \\ 
                & partitioning &     &     \\ 
\hline
\textbf{prespecified GAM} & spline  & tp{\_}mean, & tppow{\_}sprd,\\
                            & in each & tp{\_}max,  & dswrf{\_}sprd{\_}mean,\\
 & & tp{\_}mean1218 $\ast$ & tp{\_}sprd1218 $\ast$\\ 
 & & \quad cape{\_}mean1218, & \quad  cape{\_}mean1218,\\
 & & dswrf{\_}mean{\_}mean, & tcolc{\_}sprd{\_}mean,\\
 & & tcolc{\_}mean{\_}mean, & tdiff500850{\_}mean\\
 & & pwat{\_}mean{\_}mean, & \\
 & & tdiff500850{\_}mean, & \\
 & & msl{\_}diff & \\
\hline
\textbf{boosted GAM} & spline  & all & all \\
                       & in each &     &     \\
\hline
\textbf{EMOS} & linear & tp{\_}mean & tp{\_}spread \\
\hline
\end{tabular}
\caption[Table caption text]{Table of applied methods together with their type (way of including the covariates) and the included covariates for each distribution parameter.}
\label{model_specification}
%}
\end{center}
\end{table}

For the prespecified GAM model the covariates were selected manually based on meteorological knowledge and experience. Additionally, interactions are specified in advance, once between \texttt{tp{\_}mean1218} and \texttt{cape{\_}mean1218} for the location parameter and once between \texttt{tp{\_}sprd1218} and \texttt{cape{\_}mean1218} for the scale parameter. These interactions might capture events with a high potential for moderate or severe thunderstorms with large amounts of precipitation. By putting extra effort in the specification of the formulas the prespecified GAM model is provided with additional information and is therefore expected to profit from this advantage.\\
In the EMOS model the logarithmic function is chosen as the link function for the scale parameter. This choice has been made based on the results of \citealt{Gebetsberger2017} where different link functions (identity, quadratic and logarithmic function) have been compared. Moreover, testing the identity, quadratic and logarithmic function as link functions in this setting supported this choice as the logarithmic function lead to the best performance of the EMOS model in this application.\\
\\
In order to compare the performance of the chosen methods, the continuous ranked probability score (CRPS) is calculated for each model. The CRPS measures the accuracy of a probabilistic model. For one observation $y$ with the predicted distribution function $F$ the CRPS value is calculated by
$$
\text{CRPS}(y, F) = \int_{-\infty}^\infty (F(z) - \mathbf{1}(y \leq z))^2 dz
$$
where \(\mathbf{1}\) is the indicator function. \\
In this application the CRPS is evaluated out of sample on a data set that has not been used to build the model and the mean CRPS over this data set is used for validation. Since the CRPS can be seen as a generalization of the mean absolute error, a low value indicates a good fit. In applications, one big advantage in terms of interpretability is that the resulting value is on the scale of the observation. \\

>>>>

%% packages
\usepackage{amstext}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{thumbpdf,lmodern}
%\usepackage{rotating}
%\usepackage{caption}
%\captionsetup{format=hang}
%\usepackage{subcaption}
\usepackage{Sweave}
%\usepackage{enumitem}
%\usepackage{graphicx}
\usepackage{array, makecell}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{positioning,shapes,arrows,decorations.pathreplacing,calc,automata,mindmap}
%\usepackage[pagebackref=true]{hyperref}
%\usepackage{doi}

@@

\Address{
  Lisa Schlosser, Reto Stauffer, Achim Zeileis \\
  Department of Statistics \\
  Faculty of Economics and Statistics \\
  Universit\"at Innsbruck \\
  Universit\"atsstr.~15 \\
  6020 Innsbruck, Austria \\
  E-mail: \email{Lisa.Schlosser@uibk.ac.at}, \email{Reto.Stauffer@uibk.ac.at}, \email{Achim.Zeileis@R-project.org} \\
  URL: \url{https://www.uibk.ac.at/statistics/personal/schlosser-lisa/lisa-schlosser.html}, \url{http://retostauffer.org/}, \url{https://eeecon.uibk.ac.at/~zeileis/}\\

@@

@@

After explaining the novel methodology in Section~\ref{Methodology} of this paper the focus will then be put on an application in the field of probabilistic forecasting in Section~\ref{ProbabilisticForecasting}. The new method is applied on precipitation data in a mountainous region with a large number of numerical weather prediction quantities and the results are compared to those of other methods commonly used in this research area. In a first setting only one observation station is taken into consideration. The models are trained on data from 24 years and probabilistic predictions of the consecutive 4 years are validated. A brief outlook on the results of this setting is given by Figure~\ref{axams_onestation}. For the 24th of July of 2009, 2010, 2011 and 2012 the density functions of the distributions estimated by a forest model are plotted together with the corresponding observations. This illustrates that well fitting predicted expected values are provided by this model as they are close to the observed values.
Beside the expected values the predictions provided by distributional models offer a wide range of information as they specify the whole distribution and therefore enable further methods of inference and analysis. \\

@@

## get predicted parameter
# for learndata and testdata
{
  set.seed(7)
  ## disttree
  if(FALSE){
    # in sample
    pdt <- predict(res$dt, type = "parameter")
    dt_mu_l <- pdt$mu 
    dt_sigma_l <- pdt$sigma 
    pit_dt_l <- cbind(0, pnorm(res$learndata[,"robs"], mean = dt_mu_l, sd = dt_sigma_l))
    pit_dt_l[res$learndata[,"robs"]>0, 1] <- pit_dt_l[res$learndata[,"robs"]>0, 2]
  }  
  # out of sample
  pdt <- predict(res$dt, type = "parameter", newdata = res$testdata)
  dt_mu_t <- pdt$mu 
  dt_sigma_t <- pdt$sigma 
  pit_dt_t <- cbind(0, pnorm(res$testdata[,"robs"], mean = dt_mu_t, sd = dt_sigma_t))
  pit_dt_t[res$testdata[,"robs"]>0, 1] <- pit_dt_t[res$testdata[,"robs"]>0, 2]

@@

\begin{figure}[t!]
\begin{center}
<<rain_Axams_pred, fig=TRUE, echo=FALSE, width=6.5, height=4.5>>=
plot(x = x, y = y1, type = "l", col = "red", 
     main = paste0("July ", pday), ylab = "Density", 
     xlab = expression(Total~precipitation~"["~mm^(1/1.6)~"/"~"24h"~"]"),
     #xlab = expression(Total~precipitation~group("[", mm^(1/1.6)~per~day,"]")),
     #{\frac{1}{1.6}}$,
     ylim = c(0,max(y1, y2, y3, y4, pm1, pm2, pm3, pm4) + 0.01),
     xlim = c(-1.5,8))

@@

# add labels
text(x = -0.8, y = lh1, labels = "2009", col = "red", cex = 0.8)
text(x = -0.8, y = lh2, labels = "2010", col = "blue", cex = 0.8)
text(x = -0.8, y = lh3, labels = "2011", col = "darkgreen", cex = 0.8)
text(x = -0.8, y = lh4, labels = "2012", col = "purple", cex = 0.8)
@
\caption{\label{axams_onestation}Predictions of total precipitation at station Axams for July 24 in 2009, 2010, 2011 and 2012 by a distributional forest model learned on data from 1985--2008. Observations and predictions are left censored at 0. The corresponding point mass is shown at the censoring point (0).}
\end{center}
\end{figure}

%\newpage
\section{Methodology}
\label{Methodology}

@@

In our application of precipitation forecasting the response variable, which is the power transformed total precipitation amount within 24 hours, is expected to be normally distributed with mean/location parameter $\mu$ and standard deviation/scale parameter $\sigma$ and left censored with censoring point 0. (The power transformation is used to remove large parts of the skewness of the data as further explained in \ref{data}.) Therefore, the corresponding log-likelihood function with parameter vector $\theta = (\mu, \sigma)$ is

@@

The exact way of determining how and where to split varies across different tree algorithms. For distributional trees, one of two algorithms can be chosen as a framework to build the tree: either the MOB algorithm (\citealt{zeileis2008model}) or the ctree algorithm (\\citealt{hothorn2006unbiased}). Summed up, the steps of building a distributional tree using one of these algorithms are
\begin{enumerate}
\item Specify a distribution with log-likelihood function $\ell(\theta; y)$.
\item Estimate $\bold{\hat{\theta}}$ via maximum likelihood.
\item Test for associations or instabilities of the scores $s(\bold{\hat{\theta}}, y_i)$ and each partitioning variable $Z_l$.
%\item Assess whether the \emph{model scores} are associated with
%    (or change along) any of the available covariates -- e.g.,
%    using parameter instability tests (\emph{strucchange}) or
%    conditional inference (\emph{coin}).
\item Split the sample along the partitioning variable with the strongest association or instability.
    Choose the breakpoint with the highest improvement in log-likelihood.
\item Repeat steps 2--4 recursively in the subgroups
    until some stopping criterion is reached.
    %-- e.g., for significance or sample size.
%\item Choose the variable with the strongest association
%\item Choose the split point which leads to the highest improvement of the model
%\item Split and repeat 2-6 in each node until a stopping criterion is met
\end{enumerate}
The stopping criterion in step 5 can for example be a minimum number of observations in a node or a p-value for the statistical tests applied in step 3.\\
Executing these steps results in a stepwise parametric model where a complete distribution is fit in each node of the tree.\\
\\
For distributional trees the strategy of how to make a prediction for a new set of covariates $\bold{z}^{\ast} = (z_1^{\ast}, \ldots, z_m^{\ast})$ is very simple and straight forward. All observations of the learning data set $\{(y_i,\bold{z}_i)\}_{i=1\ldots,n}$ which end up in the same terminal node of the tree as the new set of covariates are selected and a distributional model is fit on the subset. To select this subset a weight vector $\bold{\omega}^{tree} = (\omega^{tree}_1, \ldots, \omega^{tree}_n)$ is included in the fitting process.
$$
\omega^{tree}_i(\bold{z}^{\ast}) = \sum_{b=1}^B I((\bold{z}_i \in \mathcal{B}_b) \land (\bold{z}^{\ast} \in \mathcal{B}_b))
$$
The predicted distribution of the new set of covariates is then fully specified by the estimated parameter $\hat{\theta}(\bold{z}^{\ast})$ where
$$
\hat{\theta}(\bold{z}^{\ast}) = \max_{\theta \in \Theta} \sum_{i=1}^n \omega^{tree}_i(\bold{z}^{\ast}) \cdot \ell(\theta; y_i).
$$

@@

\SweaveOpts{eval = TRUE}
%\newpage
%\newpage
\section{Probabilistic Forecasting of Precipitation in Complex Terrain}
\label{ProbabilisticForecasting}

@@

\subsection{Data}
\label{data}
The data set used for training and validation consists of observed daily precipitation sums provided by the National Hydrographical Service (\citealt{ehyd}) and numerical weather forecasts from the {U.S.} National Oceanic and Atmospheric Administration (NOAA).
Both, observations and weather forecasts are available from 1985 trough the year 2012, focusing on July only in this article.\\ 
Observations are available on a daily basis, each observation representing the total precipitation sum fallen within 24~hours observed at 6~UTC at 95 stations all over Tyrol and it's close surrounding. The observations are undergoing a rigorous quality check before made available to the public.  As covariates the second generation reforecast data set of the global ensemble forecast system (GEFS) is used (\citealt{hamill2013}).  This data set consists of an 11-member ensemble based on a fixed version of the numerical model and a horizontal grid-spacing of about $50~\times~50 \text{km}^2$ initialized at 0~UTC from December 1984 to present. Each of the 11 ensemble members uses slightly different perturbed initial conditions trying to predict the situation specific uncertainty of the atmospheric state.

@@

In particular, 14 basic forecast variables are considered with up to 12 variations of each of these variables being included as covariates. Since each forecast variable consists of an ensemble of individual predictions, the possible variations include the mean, the standard deviation, the minimum and the maximum over all ensemble members, noted as ensemble mean, ensemble std dev, ensemble minimum and ensemble maximum. All of these values can be calculated on the sums over 24 hours (6--6 UTC of the following day) or a 6 hour time window.\\
Another possible option is to consider the mean, minimum or maximum over 24 hours for each individual prediction and then take the ensemble mean or ensemble standard deviation, noted as ensemble mean of mean/minimal/maximal values or ensemble std dev of mean/minimal/maximal values respectively.\\
Table~\ref{covariates} shows all variables together with the number of variations and the provided variations listed in the last column. Altogether 82 covariates are available in this setting.\\
To remove large parts of the skewness of precipitation data, a power transformation has been used frequently (Box and Cox 1964). In literature, cubic (\citealt{stidd1973}) or square root (\citealt{hutchinson1998b}) transformations have often been suggested but may vary for different climatic zones or temporal aggregation periods. In \citealt{stauffer2017a} optimization methods have been applied to find the best transformation factor which resulted in a value of $\frac{1}{1.6}$. Since the data used for this optimization covers almost the same geographical region as considered in this article the same factor is chosen here for the power transformation of the response variable \texttt{tp} (total precipitation). The resulting transformed variable is expected to be normally distributed and left censored at censoring point 0. For simplicity the same transformation has been chosen for the variable \texttt{cape} (convective available potential energy).

%\newpage
\begin{table}[t!]
%\resizebox{1.4\textwidth}{!}{
\begin{minipage}{\textwidth}
\begin{tabular}{ l  c  l }
\hline
Basic covariates & {\#} & Variations\\
\hline
\textbf{tp}: total precipitation            & 12  & ensemble mean of sums over 24h, \\
\hspace*{\fill} power transformed (by $\frac{1}{1.6}$) &    & ensemble std dev of sums over 24h, \\ 
\textbf{cape}: convective                  &    & ensemble minimum of sums over 24h, \\
\hspace*{0.4cm} available potential energy &    & ensemble maximum of sums over 24h\\
\hspace*{\fill}  power transformed (by $\frac{1}{1.6}$)&    & \qquad all for 6--30 UTC\\
\hspace*{\fill}                            &    & ensemble mean of sums over 6h\\
\hspace*{\fill}                            &    & \qquad for 6--12, 12--18, 18--24, 24--30 UTC\\
\hspace*{\fill}                            &    & ensemble std dev of sums over 6h\\
\hspace*{\fill}                            &    & \qquad for 6--12, 12--18, 18--24, 24--30 UTC\\
\hline
\textbf{dswrf}: downwards short ware      & 6 & ensemble mean of mean values, \\
\hspace*{\fill} radiation flux (``sunshine'') &   & ensemble mean of minimal values,\\
\textbf{msl}: mean sea level pressure     &   &  ensemble mean of maximal values,\\
\textbf{pwat}: preciptable water          &   & ensemble std dev of mean values,\\
\textbf{tmax}: 2m maximum temperature     &   & ensemble std dev of minimal values,\\
\hspace*{1.15cm}                          &   & ensemble std dev of maximal values,\\
\textbf{tcolc}: total column-integrated   &   & \qquad all over 6--30 UTC \\
\hspace*{1.05cm} condensate               &   & \\
\textbf{t500}: temperature on 500 hPa     &   & \\
\hspace*{1cm}                             &   & \\
\textbf{t700}: temperature on 700 hPa     &   & \\
\hspace*{1cm}                             &   & \\
\textbf{t850}: temperature on 850 hPa     &   & \\
\hspace{1cm}                              &   & \\
\hline
\textbf{tdiff500850}: temperature         & 3 & ensemble mean of difference in mean,\\
\hspace*{\fill} difference 500 to 850 hPa &   & ensemble minimum of difference in mean,\\
\textbf{tdiff500700}: temperature         &   & ensemble maximum of difference in mean\\
\hspace*{\fill} difference 500 to 700 hPa &   & \qquad all over 6--30 UTC\\
\textbf{tdiff700850}: temperature         &   & \\
\hspace*{\fill} difference 700 to 850 hPa &   & \\
\hline
\textbf{msl{\_}diff}: mean sea level  & 1 & msl{\_}mean{\_}max - msl{\_}mean{\_}min\\
\hspace{1.65cm} difference            &   & \qquad over 6--30 UTC\\
\hline
\end{tabular}
\caption[Table caption text]{Basic covariates together with the number ({\#}) and the type of variations.}
\label{covariates}
\end{minipage} 
%}
\end{table}

\subsection{Models and Evaluation}
Next to distributional forests three other methods are applied to this data set and their results are compared based on CRPS values. One of the reference models is an EMOS models as this is a common choice in this type of setting / research field. The other two models are both GAMs, once in a prespecified form in terms of variable selection for each distribution parameters and once applying a boosting algorithm). 
The chosen methods and their specifications are listed in Table~\ref{model_specification}.
\begin{table}[t!]
%\resizebox{1.4\textwidth}{!}{
%\begin{minipage}{\textwidth}
\begin{center}
\begin{tabular}{ l  l   l   l }
\hline
Model & Type & Location ($\mu$) & Scale ($\log(\sigma)$) \\
\hline
\textbf{Forest} & recursive    & all & all \\ 
                & partitioning &     &     \\ 
\hline
\textbf{prespecified GAM} & spline  & tp{\_}mean, & tppow{\_}sprd,\\
                            & in each & tp{\_}max,  & dswrf{\_}sprd{\_}mean,\\
 & & tp{\_}mean1218 $\ast$ & tp{\_}sprd1218 $\ast$\\ 
 & & \quad cape{\_}mean1218, & \quad  cape{\_}mean1218,\\
 & & dswrf{\_}mean{\_}mean, & tcolc{\_}sprd{\_}mean,\\
 & & tcolc{\_}mean{\_}mean, & tdiff500850{\_}mean\\
 & & pwat{\_}mean{\_}mean, & \\
 & & tdiff500850{\_}mean, & \\
 & & msl{\_}diff & \\
\hline
\textbf{boosted GAM} & spline  & all & all \\
                       & in each &     &     \\
\hline
\textbf{EMOS} & linear & tp{\_}mean & tp{\_}spread \\
\hline
\end{tabular}
\caption[Table caption text]{Table of applied methods together with their type (way of including the covariates) and the included covariates for each distribution parameter.}
\label{model_specification}
%}
\end{center}
\end{table}

For the prespecified GAM model the covariates were selected manually based on meteorological knowledge and experience. Additionally, interactions are specified in advance, once between \texttt{tp{\_}mean1218} and \texttt{cape{\_}mean1218} for the location parameter and once between \texttt{tp{\_}sprd1218} and \texttt{cape{\_}mean1218} for the scale parameter. These interactions might capture events with a high potential for moderate or severe thunderstorms with large amounts of precipitation. By putting extra effort in the specification of the formulas the prespecified GAM model is provided with additional information and is therefore expected to profit from this advantage.\\
In the EMOS model the logarithmic function is chosen as the link function for the scale parameter. This choice has been made based on the results of \citealt{Gebetsberger2017} where different link functions (identity, quadratic and logarithmic function) have been compared. Moreover, testing the identity, quadratic and logarithmic function as link functions in this setting supported this choice as the logarithmic function lead to the best performance of the EMOS model in this application.\\
\\
In order to compare the performance of the chosen methods, the continuous ranked probability score (CRPS) is calculated for each model. The CRPS measures the accuracy of a probabilistic model. For one observation $y$ with the predicted distribution function $F$ the CRPS value is calculated by
$$
\text{CRPS}(y, F) = \int_{-\infty}^\infty (F(z) - \mathbf{1}(y \leq z))^2 dz
$$
where \(\mathbf{1}\) is the indicator function. \\
In this application the CRPS is evaluated out of sample on a data set that has not been used to build the model and the mean CRPS over this data set is used for validation. Since the CRPS can be seen as a generalization of the mean absolute error, a low value indicates a good fit. In applications, one big advantage in terms of interpretability is that the resulting value is on the scale of the observation. \\
For the purpose of comparing the performance of a distributional forest to those of other models the so-called skills score can be calculated with respect to the CRPS and one model being chosen as a reference. In that way we obtain a measure of how much a method improves the model performance in terms of CRPS values compared to the reference method.
$$
\text{SkillsScore}_{\text{method}} = 1 - \frac{\text{CRPS}_{\text{method}}}{\text{CRPS}_{\text{reference}}}
$$

%\SweaveOpts{eval = TRUE}
\subsection{Application for one station}
First of all we focus on one observation station, Axams in Tyrol. This station is selected due to its geographical closeness to Innsbruck, the capital of Tyrol. As for all other stations, daily observations of the month of July of 28 years are available from 1985 through the year 2012.\\ 
The results of fitting one distributional forest to the first 24 years of the data set and predicting the total amount of precipitation for one specific day of each of the four successive years have been shown in Figure~\ref{axams_onestation} as a motivational example for a probabilistic forecast. Staying with this fitted model we are now going to analyze variable importance of this forest model and compare its results to those of the GAM models and the EMOS model based on probability integral transform (PIT) histograms.\\
\\
As shown in Table~\ref{model_specification} the forest model includes all 82 covariates. A single tree model allows to easily analyze which covariates are used as split variables in the nodes and hence have an impact on the response variable. In a forest model though, for each of the trees a subset of these covariates is chosen randomly as possible split variables which is one of the reasons leading to different splitting variables being chosen in each tree. To get an overview of which covariates have the most influential effect on the response variable a standard measurement of variable importance for forest models has been applied. Based on the mean decrease in accuracy regarding the CRPS the 10 leading covariates of the above described forest model for station Axams learned on the first 24 year are listed in Figure~\ref{varimp}. As expected, the numerical forecast for total precipitation is the most influential variable. In particular, five variations of this variable are among the top ten of the list of covariates ordered by variable importance.
